---
title: "Forest Landscape Metrics (FLM)"
author: "IABG EO TEAM"
date: "2024"
output:
    pdf_document:
        toc: FALSE
        highlight: zenburn
    html_document:
        css: "style.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```

<!-- \newpage -->

<!-- # Results -->

```{r, echo = FALSE}
plot_files <- list.files(plotdir, "\\.tif$", full.names = TRUE)

plot_crs <- terra::rast(plot_files[1]) |> terra::crs()
aoi<- aoi |> terra::project(plot_crs)
my_adm_bound <- aoi |> sf::st_as_sf() |> st_where_is(tempdir = path$proc_dir)
adm_bound_all <- geodata::gadm(my_adm_bound, path = path$proc_dir, level = 2)
country_adm<- geodata::gadm(my_adm_bound, path = path$proc_dir, level = 0)
adm_bound <- adm_bound_all |> terra::project(terra::crs(aoi)) |> 
    terra::crop(terra::ext(aoi))

my_coltab<- data.frame(coltab(path$lc_raster |> terra::rast()))
my_coltab<- rbind(my_coltab, c(value = -9999, my_coltab[1,2:5]))


```
This section summarizes the results for the chosen AOI, located in the municipality(ies) of **`r unique(adm_bound$NAME_2)`**, in the state(s)/ of **`r unique(adm_bound$NAME_1)`**, in the country(ies) of **`r unique(adm_bound$COUNTRY[1])`**, respectively. 

```{r}
# get appropiate raster and crop it 
r <- path$lc_raster |> terra::rast()
aoi<- aoi |> terra::project(r)
r<- r |> terra::crop(aoi)

#area
area<- terra::expanse(aoi, unit = "ha", transform = TRUE) |> round(2)

# percentage forest and all other classes
r_masked<- r |> mask(aoi)
r_masked<- r_masked |> select_forest_from_glc_lcc(all = TRUE)
r_masked_forest<- r_masked |> select_forest_from_glc_lcc(binary = TRUE)
r_masked_forest_types<- r_masked |> select_forest_from_glc_lcc()

area_pixels<- terra::expanse(r_masked, unit = "ha", transform = TRUE, byValue = TRUE)
area_pixels_forest<- terra::expanse(r_masked_forest, unit = "ha", transform = TRUE, byValue = TRUE)
area_pixels_forest_type<- terra::expanse(r_masked_forest_types, unit = "ha", transform = TRUE, byValue = TRUE) |> 
    filter(value != "no_forest")

par(mfrow = c(2, 2))
world_map<- geodata::world(path = path$proc_dir)
aoi<- aoi |> terra::project(terra::crs(world_map))
plot(world_map)
plot(country_adm, add = TRUE, col = "red")
plot(country_adm)
plot(adm_bound_all[(adm_bound_all$NAME_1 %in% unique(adm_bound$NAME_1))] |> terra::aggregate(), add = TRUE, col = "red")
plot(adm_bound_all[(adm_bound_all$NAME_1 %in% unique(adm_bound$NAME_1))])

plot(aoi, add = TRUE, col = "red")
plot(r_masked)
par(mfrow = c(1,1))

```


<!-- ## Plots -->

<!-- ### Exploratory Data Analysis -->

```{r, echo = FALSE}
 forest_perc<- (100*area_pixels_forest$area[2]/(area_pixels_forest$area[2]+area_pixels_forest$area[1])) |> round(2)
forest_type_perc<- area_pixels_forest_type |> mutate(total_area = sum(area), perc = round(area/total_area,2))


```

The selected AOI has **`r area`** hectares, out of which **`r forest_perc`%** is forest. Of the forest land cover, **`r with(forest_type_perc, gsub("_"," ",paste0(glue::glue("{value} is {perc}%"), collapse = ", ")))`** The other land cover categories are **`r area_pixels$value |> paste0(collapse = ", ") |> gsub("_"," ",x=_) |> tolower()`**. In terms of area (ha), they are distributed as follows:

```{r}
par(mfrow = c(2,2))
barplot(area~value, area_pixels)
barplot(area~value, area_pixels_forest, names.arg = c("non-forest", "forest"))
barplot(area~value, area_pixels_forest_type)
par(mfrow = c(1,1))
```


`r barplot(r_masked_forest)`

`r barplot(r_masked_forest_types)`

<!-- ### Patch metrics -->

The top 5 `patch` metrics where selected for creating maps. The `patch` metrics calculated were: **`r plot_files |> fs::path_file() |> fs::path_ext_remove() |> gsub("_", " ", x = _) |> paste0(collapse = ", ")`**. `Id` is the patch id, and category is the land cover. For more information on the other metrics, go [\textcolor{blue}{here}](https://r-spatialecology.github.io/landscapemetrics/reference/index.html#patch-level-metrics).

The category map show the land cover classification used in the FLM metrics calculation. This is the land cover map from the Copernicus Global Land Servive ([\textcolor{blue}{link}](https://land.copernicus.eu/global/products/lc)). For calculating the FLM metrics, only the forest classes are used, and all other layers are treated as a single non-forest category.

```{r echo=FALSE, warning = FALSE, message = FALSE}

plot_crs <- terra::rast(plot_files[1]) |> terra::crs()
aoi<- aoi |> terra::project(plot_crs)
my_adm_bound <- aoi |> sf::st_as_sf() |> st_where_is(tempdir = path$proc_dir)
adm_bound <- geodata::gadm(my_adm_bound, path = path$proc_dir, level = 2)

adm_bound <- adm_bound |> terra::project(terra::crs(aoi)) |> 
    terra::crop(terra::ext(aoi))

my_coltab<- data.frame(coltab(path$lc_raster |> terra::rast()))
my_coltab<- rbind(my_coltab, c(value = -9999, my_coltab[1,2:5]))

par(mfrow = c(2,2))
for (i in 1:length(plot_files)) {
    tmp_name <- fs::path_file(plot_files[i]) |> 
        fs::path_ext_remove()
    tmp <- terra::rast(plot_files[i])
    if(tmp_name == "category") coltab(tmp) <- my_coltab
    plot(tmp,main = tmp_name)
    lines(aoi, col = "red", alpha = 0.6)
    lines(adm_bound, lwd = 2.5, alpha = 0.5)
}
par(mfrow = c(1,1))

```

<!-- # General Information -->

The following explanation of the metrics has been widely reused from the following source: [\textcolor{blue}{Original}](https://r-spatialecology.github.io/landscapemetrics/articles/general_background.html#levels-of-landscape-metrics)


Ecology and landscape ecology mainly studies interactions between organisms and their environment. In a wider sense, the environment perceived by the organism can be defined as a landscape. This can spread from huge areas for, e.g., mobile, large-bodied mammals to much smaller areas for, e.g., immobile insects and always depends on the research question.

Landscape metrics are tools to characterize a landscape. This includes mainly describing the composition and configuration of a landscape. While the composition basically describes how much of the landscape is occupied by a certain land cover type, the configuration mainly describes the spatial arrangement of the land cover types. The basic idea of landscape metrics is to condense as much information as possible into single number.

<!-- ## Levels of metrics -->

**Landscape metrics** can be calculated for three different levels (or "scales"). Each level contains information about different aspects of the landscape. Level of interest depends largely on the research question, and often a combination of several levels can be useful. Similar, also the decision of which metric to use depends on the research question asked.

**Patch level metrics**. A patch is defined as neighboring cells belonging to the same class i. Hereby, the metrics use the 8-neighbors rule (Queen's case) to identify patches. Patch level metrics are calculated for every patch in the landscape, regardless of the class that the patch belongs to. The output will match the number of patches present in the landscape. These metrics are also often the basis for metrics of the other two levels.

**Class level metrics** summarizes all patches belonging to one class i. These metrics can be either the "distribution" of patch level metrics of all patches of class i (e.g., the mean) or consider only patches of class i for the calculations of the metric. Regardless of the mathematical background, the output will always match the number of classes present. Class level metrics are suitable to describe the composition and configuration of the landscape.

**Landscape level metrics** summarizes the whole landscape into one value. This can either be done by summarizing metrics of lower levels or calculating a metric including all patches and classes. Following, the output will always be just one number. Landscape level metrics are suitable for condensing information about the landscape into just one value.

<!-- ## Classes of landscape metrics -->

There are six different classes of landscape metrics implemented:

1. **Area and edge metrics** describe the size of patches and classes and the amount of edge. An edge is defined as the border between two patches of class i and k in meters. These metrics mainly characterize the composition of the landscape and are able to show dominance or rareness of classes.

2. **Shape metrics** describe the shape of patches, mainly by using its area and perimeter. This can be important for many research questions, because, e.g., even though, being equal in size, long and narrow patches have probably different characteristics than a squared patch of the same size.

3. **Core metrics** describe the area of patches that are not an edge. These metrics can be interesting for research questions in which, e.g., only areas that are not influenced by neighboring patches of a different class are of interest.

4. **Aggregation metrics** describe if patches (of the same class) are rather clumped (aggregated) or tend to be isolated. Following, these metrics describe mainly the spatial configuration of the landscape.

5. **Diversity metrics** are only available on the landscape level. They describe the abundance and dominance/rareness of classes. Thereby, they show the diversity of present classes.

6. **Complexity metrics** For more information, please see [\textcolor{blue}{New Metrics outside FRAGSTATS}](https://r-spatialecology.github.io/landscapemetrics/articles/new_metrics.html).
