---
title: "Landscape Metrics"
author: "Emin"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
#Thu Nov 16 12:25:40 2023----------
# D. Emin -------------------------

```

## Introduction

Write short ***intro*** on what landscape metrics are from the theory section.

form J.

## Inputs

[link here](www.google.de)

The input for this analysis comes from Hessen. We will use the Forest Data and a set of municipalities (Gemeinde) simulating an use case in which the Gemeinde wants to get some useful statistics about their forest.

```{r}
library(sf)
library(gdfR)
library(tidyverse)
library(landscapemetrics)
library(fs)
library(terra)
library(leaflet)
library(htmlTable)
```

```{r}
#inputs
aoi <- st_read('data/aoi_gadm_test.shp')
landscape <- rast('data/hessen_type_test2.tif')

```

The complementary data to run these scripts is located in

`\\gdfs06\DATEN09\ESA-EO4SDG_D9\02_analysis\FLM\Hessen\test_data\gitlab_v1`.

To run the example, clone the repository and copy the `data` folder located in the location above into the project folder. The final project folder structure should look like this:

    .
    +-- data
    |   +-- aoi
    |   |   +-- aoi_gadm_test.cpg
    |   |   +-- aoi_gadm_test.dbf
    |   |   +-- aoi_gadm_test.prj
    |   |   +-- aoi_gadm_test.sbn
    |   |   +-- aoi_gadm_test.sbx
    |   |   +-- aoi_gadm_test.shp
    |   |   +-- aoi_gadm_test.shx
    |   |   +-- test_aoi2.cpg
    |   |   +-- test_aoi2.dbf
    |   |   +-- test_aoi2.prj
    |   |   +-- test_aoi2.sbn
    |   |   +-- test_aoi2.sbx
    |   |   +-- test_aoi2.shp
    |   |   \-- test_aoi2.shx
    |   \-- hessen
    |       +-- hessen_type_test.tif
    |       +-- hessen_type_test2.tif
    |       \-- metrics.csv
    +-- eo4sdg-forest-flm.Rproj
    +-- R
    |   +-- 00-inputs.R
    |   +-- 01-aoicheck.R
    |   +-- 02-clip.R
    |   \-- 03-flm.R
    \-- README.md

## Overview of R-package `landscapemetrics`

Here shortly review the package itself, how it s build etc. put some useful links.

Checking the validity of a landscape.

```{r}
# convert to factor (for plotting purposes)
cls <- data.frame(id=1:3, cover=c("Nadelwald","Laubwald", "Mischwald"))
levels(landscape) <- cls
plot(landscape)
plot(st_geometry(aoi), add = T)
text(st_coordinates(st_centroid(aoi)), aoi$NAME_4)
```

```{r}
check_landscape(landscape)
```

```{r}
show_patches(landscape)
```

We can check a specific metric.

```{r}
show_lsm(landscape, what = 'lsm_p_area')
```

```{r}
# calculate the area metrics (and some more)
area_metrics <- 
    sample_lsm(landscape,
           aoi,
          metric = c('area', "ca",  'ta', 'pland',
                     'core', 'tca','cpland', 'cai', 'ed', 'lpi', 'te'),
          plot_id = aoi$NAME_4
          ) %>% 
    mutate(class = recode(class, !!!setNames(cls$cover, cls$id))) 
```

Summary for a single Gemeinde.

```{r}
area_metrics %>% 
    filter(plot_id == 'Alheim', 
           level == "class") %>% 
    pivot_wider(id_cols = metric, names_from = class, values_from = value) %>% 
    mutate_if(is.numeric, ~round(.,2)) %>% 
    knitr::kable()
```

```{r}
area_metrics %>% 
    filter(level == 'landscape') %>% 
    pivot_wider(id_cols = metric, names_from = plot_id, values_from = value) %>% 
    mutate_if(is.numeric, ~round(.,2)) %>% 
    knitr::kable()
```

```{r}
area_metrics_w <- 
    area_metrics %>% 
    filter(level != 'patch') %>% 
    pivot_wider(id_cols = plot_id,
                names_from = c("level", 'class', 'metric'),
                values_from = value)

DT::datatable(area_metrics_w)
```

## Exploratory data analysis

Here we will perform a first analysis into the behaviour of the different metrics.

For this examples we will use a precalculated `.csv` containing all metrics.

```{r}
metrics <- read_csv("data/metrics.csv")
```

### Plotting Functions

We defined several standardized ways to plot the results.

```{r}
my_ggplot <- function(data, x, y, title, subtitle=NULL){
    ggplot(data) +
        ggtitle(title, subtitle) +
        geom_bar(aes({{x}},
                     {{y}},
                     fill=class),
                 stat="identity",
                 width=.5,
                 #position = "dodge"
        ) +
        scale_x_discrete(guide = guide_axis(angle = 30))
}


my_boxplot <- function(data, means, .metric, title, subtitle = NULL){

    ggplot(data = filter(data,
                         metric == .metric),
           aes(x    = class,
               y    = value,
               fill = class)) +
        geom_boxplot() +
        geom_line(data  = filter(means, metric == .metric),
                  aes(x = class,
                      y = value,
                      group = interaction(plot_id, metric))) +
        facet_wrap(~plot_id,
                   scales = "free") +
        ggtitle(str_to_title(title), subtitle) +
        theme(legend.position = "none")
}
```

### Area and edge metrics

Note: we consider each administrative unit (9 in total) as a landscape

Overview of landscape metrics grouped by type `lm_by_type <- list_lsm() |> group_by(type) |> nest()`

#### Edge {#sec-edge}

| Metric       | Patch | Class | Landscape |
|--------------|-------|-------|-----------|
| TOTAL EDGE   |       | ✔️    | ✔️        |
| EDGE DENSITY |       | ✔️    | ✔️        |

#### Area {#sec-area}

| Metric              | Patch | Class | Landscape |
|---------------------|-------|-------|-----------|
| AREA                | ✔️    | ✔️    | ✔️        |
| PLAND               |       | ✔️    |           |
| CORE AREA           | ✔️    | ✔️    | ✔️        |
| CPLAND              |       | ✔️    |           |
| CORE AREA INDEX     |       |       |           |
| LARGEST PATCH INDEX |       | ✔️    | ✔️        |

```{r}
# Area and edge metrics --------------------------------------------------

#Copy the sample_lsm output as xx and group by landscape extent name
metrics_a <-
    metrics[[1]] |>
    group_by(plot_id)

#Also possible to nest the dataframe so that we have all sample_lsm results categorised by landscape extent
# xx <- x[[1]] |>
#     group_by(plot_id) |>
#     nest()

# Calculate the area for each "landscape" out of the aoi data
aoi_w_area <-
    aoi |>
    mutate(plot_id = NAME_4) |>
    select(plot_id) |>
    mutate(area = units::set_units(st_area(geometry), "ha")) |>
    st_drop_geometry()


# Do a left join to the dataset with total landscape area and define it as a new variable
metrics_a <-
    metrics_a |>
    left_join(aoi_w_area) |>
    mutate(plot_id,
           area             = as.numeric(area),
           relative_area    = (value/area)*100,
           class            = factor(class))

#------------ # Mean patch size for a class --------------------------------
# The metric summarises each class as the mean of all patch areas belonging to class i.
# The metric is a simple way to describe the composition of the landscape.
# Especially together with the total class area (lsm_c_ca),
# it can also give an an idea of patch structure (e.g. many small patches vs. few larges patches).


# Tried boxplots on patch level, but too many outliers.
# Boxplots are very squished

xx_mean_area_class <-
    metrics_a |>
    filter(level == "class", str_detect(metric, "_mn")) |>
    mutate(metric = str_replace(metric, "_mn", ""))


plot_mean_patch_size <-
    metrics_a |>
    filter(level  == "patch") |>
    my_boxplot(means    = xx_mean_area_class,
               .metric  = "area",
               title    = "Mean size",
               subtitle = "Patch level" )

# Bar plots on class level
plot_size_class <-
    metrics_a |>
    filter(metric == "area_mn",
           level  == "class") |>
    my_ggplot(plot_id,
              value,
              "Mean size",
              "Class level - unclassified area not included" )

#
#------------ # Total class area --------------------------------
# xx_totalarea_class <-
#     metrics_a |>
#     filter(metric == "ca") |>
#     mutate(total_area = value) |>
#     my_ggplot(plot_id,
#               total_area,
#               "Total area",
#               "Class level - unclassified area not included" )



#------------ # Percentage of landscape area -------------------------------
# It is the percentage of the landscape belonging to class i.
# It is a measure of composition and because of the relative character directly comparable
# among landscapes with different total areas.

##
##NOTE: We include two pland calculations in this chapter. One with background ("manual" calculation)
##      and one without (built-in metric)

# Use the total landscape area to calculate class area percentage (background considered)
plot_pland_w_backg <-
    metrics_a |>
    filter(metric == "ca") |>
    my_ggplot(plot_id,
              relative_area,
              title     = "Percentage of landscape",
              subtitle  = "Class level - unclassified area included" )



# Calculate percentage of landscape per class using the built-in metric pland (background not considered)
# xx_pland_wout_backg  <-
#     metrics_a |>
#     filter(metric == "pland",
#            level  == "class") |>
#     mutate(percentage_of_land = value) |>
#     my_ggplot(landscape_extent,
#               percentage_of_land,
#               "Percentage of Landscape",
#               "Class level - unclassified area not included")



#------------ # Radius of gyration (class level) #---------------------------------------------
# Radius of gyration is a measure of patch extent (i.e., how far-reaching it is);
# thus, it is effected by both patch size and patch compaction.
# Note that the choice of the 4-neighbor or 8-neighbor rule for delineating patches
# will have an impact on this metric.
# Behaviour: Approaches GYRATE_MN = 0 if every patch is a single cell. Increases, without limit, when only one patch is present.


plot_gyrate <-
    metrics_a |>
    filter(metric == "gyrate_mn",
           level  == "class") |>
    my_ggplot(plot_id,
              value,
              title     = "Gyration value",
              subtitle  = "Class level")




#------------ # Largest patch index #----------------------------------
# It is the percentage of the landscape covered by the corresponding largest patch of each class i.
# It is a simple measure of dominance.

plot_lpi <-
    metrics_a |>
    filter(metric == "lpi",
           level  == "class") |>
    my_ggplot(plot_id,
              value,
              title     = "Largest Patch Index",
              subtitle  = "Class level")



#------------ # Edge density #----------------------------------------------------------------
# The edge density equals the sum of all edges of class i in relation to the landscape area.
# The boundary of the landscape is only included in the corresponding total class edge length
# if count_boundary = TRUE. The metric describes the configuration of the landscape,
# e.g. because an aggregation of the same class will result in a low edge density.
# The metric is standardized to the total landscape area, and therefore comparisons among
# landscapes with different total areas are possible.


plot_ed <-
    metrics_a |>
    filter(metric == "ed",
           level  == "class") |>
    my_ggplot(plot_id,
              value,
              title     = "Edge density",
              subtitle  = "Class level")









# Shape metrics ----------------------------------------------------------

#Copy the sample_lsm output as xx and group by landscape extent name
xx_shape <- x_shape[[1]] |>
    group_by(plot_id)

# Calculate the area for each "landscape" out of the aoi data
aoi_w_area <- aoi |>
    mutate(plot_id = NAME_4) |>
    select(plot_id) |>
    mutate(area = units::set_units(st_area(geometry), "ha")) |>
    st_drop_geometry()


# Do a left join to the dataset with total landscape area and define it as a new variable
xx_shape <- xx_shape |>
    left_join(aoi_w_area) |>
    mutate(plot_id,
           area             = as.numeric(area),
           relative_area    = (value/area)*100,
           class            = factor(class))




#-------------# Shape index #-------------------------------------------------
# It describes the ratio between the actual perimeter of the patch and the
# square root of patch area. It is a simple measure of shape complexity.
#
# Behavior: Equals SHAPE_MN = 1 if all patches are squares. Increases,
# without limit, as the shapes of patches become more complex.
#

# Creating boxplots to visualize metrics per class and compare them between landscapes


# To plot a line connecting all mean values between classes (regardless of metric and landscape)
# Extract the mean values of each metric per class and delete the "_mn" suffix
xx_mean_class <-
    xx_shape |>
    filter(level == "class",
           str_detect(metric, "_mn")) |>
    mutate(metric = str_replace(metric, "_mn", ""))

#Check for correct replacement using
#unique(xx_mean_class$metric)



xx_shapeindex <-
    xx_shape |>
    filter(level  == "patch") |>
    my_boxplot(means = xx_mean_class,
               .metric = "shape",
               title = "shape index",
               subtitle = "Patch level")



#-------------# Related circumscribing circle #-------------------------------------------------
# The metric characterises the compactness of the patch and is comparable among patches with different area.
# Behaviour: CIRCLE = 0 for a circular patch and approaches CIRCLE = 1 for a linear patch.


xx_circle <-
    xx_shape |>
    filter(level  == "patch") |>
    my_boxplot(means = xx_mean_class,
               .metric = "circle",
               title= "Related circumscribing circle",
               subtitle = "Patch level")


#-------------# Contiguity index #-------------------------------------------------
# The metric characterises the compactness of the patch and is comparable among patches with different area.
# Behaviour: CIRCLE = 0 for a circular patch and approaches CIRCLE = 1 for a linear patch.


xx_contig <-
    xx_shape |>
    filter(level  == "patch") |>
    my_boxplot(means = xx_mean_class,
               .metric = "contig",
               title= "Contiguity index",
               subtitle = "Patch level")








# Core metrics ----------------------------------------------------------


#Copy the sample_lsm output as xx and group by landscape extent name
xx_core <- x_core[[1]] |>
    group_by(plot_id)

# Calculate the area for each "landscape" out of the aoi data
aoi_w_area <- aoi |>
    mutate(plot_id = NAME_4) |>
    select(plot_id) |>
    mutate(area = units::set_units(st_area(geometry), "ha")) |>
    st_drop_geometry()


# Do a left join to the dataset with total landscape area and define it as a new variable
xx_core <- xx_core |>
    left_join(aoi_w_area) |>
    mutate(plot_id,
           area             = as.numeric(area),
           relative_area    = (value/area)*100,
           class            = factor(class))


#-------------# Core area index #-------------
# Range: 0 <= CAI_MN <= 100
# Behaviour: CAI_MN = 0 when all patches have no core area and approaches
#            CAI_MN = 100 with increasing percentage of core area within patches.


# Creating boxplots to visualize metrics per class and compare them between landscapes


# To plot a line connecting all mean values between classes (regardless of metric and landscape)
# Extract the mean values of each metric per class and delete the "_mn" suffix
xx_mean_class_core <-
    xx_core |>
    filter(level == "class",
           str_detect(metric, "_mn")) |>
    mutate(metric = str_replace(metric, "_mn", ""))

#Check for correct replacement using
unique(xx_mean_class_core$metric)

xx_coreindex <-
    xx_core |>
    filter(level  == "patch") |>
    my_boxplot(means = xx_mean_class_core,
               .metric = "cai",
               title = "core area index",
               subtitle = "Patch level")



#-------------# Core percentage of land (not yet fully done) #-------------
# It is the percentage of core area of class i in relation to the total landscape area.
# A cell is defined as core area if the cell has no neighbour with a different value than
# itself (rook's case). Because CPLAND is a relative measure,
# it is comparable among landscapes with different total areas.

# Units; Percentage
# Range: 0 <= CPLAND < 100
# Behaviour: Approaches CPLAND = 0 if CORE = 0 for all patches.
#            Increases as the amount of core area increases, i.e. patches become
#            larger while being rather simple in shape.

xx_cpland_w_backg <-
    xx_core |>
    filter(metric == "cpland") |>
    my_ggplot(plot_id,
              relative_area,
              "Core area percentage of landscape",
              "Class level - unclassified area included" )



# # Calculate core area percentage of landscape per class using the built-in metric cpland (background not considered)
xx_cpland_wout_backg  <-
    xx_core |>
    filter(metric == "cpland",
           level  == "class")|>
    my_ggplot(plot_id,
              value,
              "Core area percentage of landscape",
              "Class level - unclassified area not included")









# TODO
# --------------------------------# Visualisation # -------------------------------
# This section takes the output from our calculated *shape metrics* as an example.
# You need to run the section "Shape metrics" in this file before continuing with the visualisation


# Create a single boxplot per metric per landscape (region)
my_single_boxplot <- function(data, land, .metric, title, subtitle = NULL){
    data |>
        filter(level  == "patch",
               plot_id == land,
               metric == .metric) |>
        ggplot(aes(x = class,
                   y = value,
                   fill = class)) +
        geom_boxplot() +
        facet_wrap(~plot_id,
                   scales = "free") +
        ggtitle(str_to_title(title), subtitle) +
        theme(legend.position = "none")
}



xx_shape_alheim <-
    my_single_boxplot(data = xx_shape,
                      land = "Alheim",
                      #means = xx_mean_class,
                      .metric = "shape",
                      title = "shape index",
                      subtitle = "Alheim")

#
# xx_shape_al <-
#     xx2 |>
#     filter(level  == "patch",
#            plot_id == "Alheim",
#            metric == "shape") |>
#     ggplot(aes(x = class,
#                y = value,
#                fill = class)) +
#     geom_boxplot() +
#     geom_line(data = filter(means, metric == .metric),
#               aes(x = class,
#                   y = value,
#                   group = interaction(plot_id, metric))) +
#     facet_wrap(~plot_id,
#                scales = "free") +
#     ggtitle(str_to_title("shape index"), "Alheim") +
#     theme(legend.position = "none")
```

## Leaflet Proof of Concept

Here we show how the final product might look like on the platform. Click on a single municipality to get some useful stats about the forest.

```{r}
# rejoin with the aoi
area_metric_spatial <- left_join(aoi, area_metrics_w, by = c("NAME_4" = "plot_id"))

#select some stats to visualise on an interactive map
content <- list()
for (i in 1:nrow(area_metric_spatial)){
    content[[i]] <- 
        st_drop_geometry(area_metric_spatial)[i,c(40:53)] %>%
        pivot_longer(everything()) %>% 
        mutate(value = round(value, 2))
}

# convert them to html strings (expected by leaflet)
content2 <- map(content, htmlTable) %>% unlist() 
```

```{r, fig.height=12, fig.width=12}
# proof of concept using leaflet
area_metric_spatial %>%
    st_transform(4326) %>%
    leaflet() %>%
    addTiles() %>%
    addPolygons(
        color = "#444444",
        weight = 1,
        smoothFactor = 0.5,
        opacity = 1.0,
        fillOpacity = 0.5,
        fillColor = ~ colorQuantile("YlOrRd", landscape_NA_ta)(landscape_NA_ta),
        highlightOptions = highlightOptions(
            color = "white",
            weight = 2,
            bringToFront = TRUE
        ),
        popup = content2
    )

```

```{r}
# Andres
aoi |>
    st_transform(crs = 4326) |>
    leaflet() |>
    addTiles() |>
    addPolygons(color = "#444444",
                weight = 1,
                smoothFactor = 0.5,
                opacity = 1.0,
                fillOpacity = 0.5,
                highlightOptions = highlightOptions(color = "white",
                                                    weight = 2,
                                                    bringToFront = TRUE),
                popup = popupGraph(xx_shape_alheim, width = 300, height = 300)
    )




```

Source for adding graphs to maps with leaflet: <https://github.com/r-spatial/leafpop> alternative to what I used: addPopupGraphs(list(xx_shape_alheim), group = "xxx", width = 300, height = 400)
